---
title: "Hello World - compilation details"
author: "mikefc@coolbutuseless.com"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
  library(r64)  
})

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r}
library(dplyr)
library(r64)

asm <- '
*=$0801
  .byte $0c, $08, $0a, $00, $9e, $20
  .byte $32, $30, $38, $30, $00, $00
  .byte $00

*=$0820
      ldx #$00
loop  lda message,x
      and #$3f
      sta $0400,x
      inx
      cpx #$0c
      bne loop

      rts

message
    .text "Hello World!"
'
```


r64 defines a set of tokens in the ASM file
------------------------------------------------------------------------------

```{r}
#-----------------------------------------------------------------------------
#' Regex patterns for parsing 6502 assembly
#-----------------------------------------------------------------------------
asm_patterns <- c(
  newline    = '\n',
  whitespace = '\\s+',
  PC         = '\\*',
  immediate  = '#\\$[0-9a-fA-F]{2}',
  word       = '\\$[0-9a-fA-F]{4}',
  byte       = '\\$[0-9a-fA-F]{2}',
  opcode     = "\\b(ADC|AHX|ALR|ANC|AND|ARR|ASL|AXS|BCC|BCS|BEQ|BIT|BMI|BNE|BPL|BRK|BVC|BVS|CLC|CLD|CLI|CLV|CMP|CPX|CPY|DCP|DEC|DEX|DEY|EOR|INC|INX|INY|ISC|JMP|JSR|LAS|LAX|LAX|LDA|LDX|LDY|LSR|NOP|ORA|PHA|PHP|PLA|PLP|RLA|ROL|ROR|RRA|RTI|RTS|SAX|SBC|SEC|SED|SEI|SHX|SHY|SLO|SRE|STA|STX|STY|TAS|TAX|TAY|TSX|TXA|TXS|TYA|XAA|adc|ahx|alr|anc|and|arr|asl|axs|bcc|bcs|beq|bit|bmi|bne|bpl|brk|bvc|bvs|clc|cld|cli|clv|cmp|cpx|cpy|dcp|dec|dex|dey|eor|inc|inx|iny|isc|jmp|jsr|las|lax|lax|lda|ldx|ldy|lsr|nop|ora|pha|php|pla|plp|rla|rol|ror|rra|rti|rts|sax|sbc|sec|sed|sei|shx|shy|slo|sre|sta|stx|sty|tas|tax|tay|tsx|txa|txs|tya|xaa)\\b",
  byte_inst  = '\\.byte',
  text_inst  = '\\.text',
  rtext_inst = '\\.rtext',
  rbyte_inst = '\\.rbyte',
  lbracket   = '\\(',
  rbracket   = '\\)',
  langle     = '<',
  rangle     = '>',
  text       = '".*?"',
  comma      = ",",
  colon      = ":",
  equals     = '=',
  comment    = '(;[^\n]*)',
  x          = '(x|X)',
  y          = '(y|Y)',
  symbol     = '[^\\s:,]+'
)
```




Each line is split into tokens
------------------------------------------------------------------------------

* Each line is split into tokens
* any `comma`, `whitespace` or `comment` tokens are discarded
* any blank lines are discarded

```{r}
line_tokens <- r64::create_line_tokens(asm)
line_tokens
```


The main `prg_df` datastructure is created from `line_tokens`
------------------------------------------------------------------------------

```{r}
prg_df <- r64::create_prg_df(line_tokens)
```

```{r echo=FALSE}
prg_df %>% 
  select(-argbytes, -value, -op, -opcommand, -nargbytes, -nopbytes) %>%
  knitr::kable()
```


Calculate addresses and replace any symbols with these values
------------------------------------------------------------------------------

```{r}
prg_df <- r64::process_xrefs(prg_df)
```

```{r echo=FALSE}
prg_df %>% 
  select(-argbytes, -hexbytes, -value, -op, -symbolop, -nopbytes, -nargbytes, -opcommand, -init_addr) %>% 
  knitr::kable()
```






Add zero padding so there are no gaps
------------------------------------------------------------------------------

* PRG files must represent a contiguous sequence of bytes - no gaps allowed!
* Find and insert sequences of zeros to ensure all gaps are filled.

```{r}
prg_df <- r64::process_zero_padding(prg_df)
```

```{r echo=FALSE}
prg_df %>% 
  select(-argbytes, -hexbytes, -value, -op, -symbolop, -nopbytes, -nargbytes, -opcommand, -init_addr) %>% 
  knitr::kable()
```


Extract the bytes from `prg_df`
------------------------------------------------------------------------------

* `prg_df$bytes` is a list column
* Use `purrr::flatten()` to convert to an integer vector (and remove any NULL entries)

```{r}
prg_df$bytes %>%
  purrr::flatten_int() %>%
  as.raw()
```




Add the loading address to get the complete PRG
------------------------------------------------------------------------------

* PRG files are prefixed by their first address so they can be loaded in the 
  right location.
* Take the first address in `prg_df` and convert it to 2 bytes in low-byte/high-byte
  format using `r64::address_to_bytes()`


```{r}
# The following is equivalent to:  r64::extract_prg_bytes(prg_df)
as.raw(c(address_to_bytes(prg_df$addr[1]), purrr::flatten_int(prg_df$bytes)))
```


Save the PRG
------------------------------------------------------------------------------

* Just write the bytes directly to file with `writeBin()` - no other processing is needed.














